---
title: "Coding Sample MWDO Application"
author: "Helena"
date: "2026-04-09"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

---
editor_options: 
  markdown: 
    wrap: 72
---

# Accessing Survey Data for POLTRACK

This document explains and demonstrates a variety of approaches
accessing the survey data for all waves of the POLTRACK project.
Documentation for the database can be found
[here](https://docs.google.com/document/d/1_pZkTmcTeoQhNqj_iO9wo1oK0_F_mRg2Qg7mTIKGyOI/edit#heading=h.8tvak299rjge "Survey Documentation"),
along with keys for the meaning of nominal variables.

### Connecting to database

Packages and password installation instructions can be found
[here](https://github.com/lisamerten/poltrack#readme "Cornelius Poltrack Markdown")

```{r}
library(DBI)
library(tidyverse)

con_survey <- dbConnect(RMySQL::MySQL(), 
                        dbname = "####".
                        host = '####', 
                        port = "####", 
                        user = '####',
                        password = Sys.getenv("password"))
dbListTables(con_survey)
```

Be sure to disconnect from the server when you are done.

```{r}
dbDisconnect(con_survey) 
```

### Query Options

The following are examples of how to query the data set for various
combinations of variables and waves. These are the direct SQL queries
needed to extract the data as a wide format table. There is also a query
generator, if you would like to avoid the SQL available here --\>
[Multiple Waves - Query Generator](#multiple-waves---query-generator).

#### One Wave

For selecting one variable in the output structure ID/Question (wide
format table) , using the variable partyid for demonstration purposes.
The code here also uses wave 1, to extract from any other specific wave
this should be adjusted directly.

1.  One variable

    ```{r}
    dbGetQuery(con_survey, "SELECT p_0002, 
          MAX(CASE WHEN Question = 'partyid' THEN Answer END) AS partyid
          FROM surveys_pooled
          WHERE Wave = 1
          GROUP BY p_0002")
    ```

2.  Multiple variables

    Using partyid and edu as an example. Two variations are shown for
    querying multiple variables:

    ```{r}
    dbGetQuery(con_survey, "SELECT p_0002, 
          MAX(CASE WHEN Question = 'affective_groups_1' THEN Answer END) AS affective_groups_1,
          MAX(CASE WHEN Question = 'edu' THEN Answer END) AS edu,
          MAX(CASE WHEN Question = 'age' THEN Answer END) AS age
          FROM surveys_pooled
          WHERE Wave = 1
          GROUP BY p_0002")
    ```

3.  All Variables

    ```{r}
    all_variables <- dbGetQuery(con_survey, "SELECT DISTINCT Question FROM surveys_pooled")

    # Create the SQL query dynamically to retrieve the maximum value for all variables
    query <- paste(
      "SELECT p_0002,", 
      paste(
        sapply(all_variables$Question, function(var) {
          paste(
            "MAX(CASE WHEN Question = '", var, "' THEN Answer END) AS ", var, sep = ""
          )
        }),
        collapse = ", "
      ),
      "FROM surveys_pooled WHERE Wave = 1 GROUP BY p_0002",
      sep = " "
    )

    # Execute the query
    result <- dbGetQuery(con_survey, query)
    ```

#### Multiple Waves - Query Generator {#multiple-waves---query-generator}

The query generator is a function which aromatically queries and
re-formats the data from any specified combination of questions and
waves. The result is a wide format table. [Function
Code](https://github.com/lisamerten/poltrack/blob/main/preprocessing/SQL_query_function.R)

```{r}
 
source(#YOUR PATH#)

output <- SQL_query_function(variables, waves, homogenize)
```

-   Variables \<- list of variables

-   Waves \<- list of wave numbers

-   Homogenize \<- a command of "YES" or "NO" which determines whether
    variables that are not repeatedly collected for individual
    participants are combined into one column. Advantage is all are
    neatly in one column, disadvantage being that their collection wave
    for that participant is lost.

Example:

```{r}

## Specify the variables
variables_to_query <- c('sonntagsfrage', 'age', 'edu')

## Specify Waves
waves <- c(1, 2)

## Apply function
output <- SQL_query_function(variables_to_query, waves, "YES")
```

### Notes

Issue with format / values of survey variables such as edu

-   The variable "edu" is a factor variable when read from the db, which
    should be transformed once to numeric and then back to factor before
    recoding.
-   This is because in the original form of the variable there are
    several values for the same value (e.g. 8 and 8.0), which can lead
    to values being forgotten when recoding. The problem is solved by
    converting to numeric and back to factor.

#### Additonal information Multiple Waves - Query Generator Code

```{r}

SQL_query_function <- function(variables, waves, homogenize = "YES") {
  
  single_wave_vars <- c(
    "einwilligung", "polinterest", "newsinterest", "partyid", "leftright",
    "partyid_icons_text", "party_importance", "party_pronouns",
    "bundesland", "edu", "nationality", "inctotal",
    "age", "gender", "ard_aspration", "zdf_aspration",
    "bild_aspration", "sz_aspration", "faz_aspration", "spiegel_aspration",
    "stern_aspration", "zeit_aspration", "focus_aspration", "welt_aspration",
    "ntv_aspration", "taz_aspration", "dritten_aspration"
  )
  
  # Initialize empty result
  result <- NULL
  
  # Loop through waves
  for (wave in waves) {
    query <- paste0(
      "SELECT id_survey, ",
      paste(
        sapply(variables, function(var) {
          paste0("MAX(CASE WHEN Question = '", var, "' THEN Answer END) AS ", var)
        }),
        collapse = ", "
      ),
      " FROM surveys_pooled WHERE Wave = ", wave, " GROUP BY id_survey"
    )
    
    wave_data <- dbGetQuery(con_survey, query)
    
    # Add wave suffix if more than one wave
    if (length(waves) > 1) {
      colnames(wave_data)[-which(colnames(wave_data) == "id_survey")] <-
        paste0(colnames(wave_data)[-which(colnames(wave_data) == "id_survey")], "_W", wave)
    }
    
    # Merge with previous results
    if (is.null(result)) {
      result <- wave_data
    } else {
      result <- merge(result, wave_data, by = "id_survey", all = TRUE)
    }
  }
  
  # Homogenize if requested
  if (homogenize == "YES") {
    
    for (var in single_wave_vars) {
      # Find columns for this variable across waves
      matching_cols <- grep(paste0("^", var, "_W[0-9]+$"), names(result), value = TRUE)
      
      if (length(matching_cols) > 0) {
        # Replace invalid codes with NA
        result[matching_cols] <- lapply(result[matching_cols], function(col) {
          ifelse(col %in% c(-77, -66, -99, "-77", "-66", "-99", "-77.0", "-66.0", "-99.0"), NA, col)
        })
        
        # Coalesce into a single column
        result[[var]] <- do.call(dplyr::coalesce, result[matching_cols])
        
        # Remove original wave columns
        result <- result[, !(names(result) %in% matching_cols)]
      }
    }
  }
  
  return(result)
}
```
